manuscript

### 2025-01-20 doc/Manuscript.qmd
begin introduction

---
title: "manuscript"
format: html
editor: visual
---

@@

>>>>

---
format: pdf
bibliography: references.bib
---

@@

Amphibians are the most threatened class of vertebrates. Disease, habitat loss, and, increasingly, climate change drive their global decline [@luedtke2023]. The endangered mountain yellow-legged frog species complex (*Rana muscosa* and *R. sierrae*) is emblematic of such declines. **Citation**. These frogs are endemic to the Sierra Nevada and Transverse mountain ranges of California and Nevada, and are characterized by \_\_\_\_\_\_. Non-native fish introductions across their range have severely reduced suitable habitat for the species[@knapp2000], and the fungal pathogen *Batrachochytrium dendrobatadis* (Bd) causes mass mortality and rapid population declines [@vredenburg2010]. As a result, more than 90% of *R. muscosa* and *R. sierrae* populations have been extirpated from their historical range [@vredenburg2007].

Despite these substantial losses, mountain yellow-legged frogs provide rare examples of amphibian resilience in the face of multiple stressors. Following the removal of non-native fish from alpine lakes, *R. muscosa* population density can increase significantly [@vredenburg2004][@knapp2007]. Additionally, some mountain yellow-legged frog populations exhibit resistance to Bd infection and naturally recover despite the pathogen's continued presence in the environment [@knapp2016]. These Bd-persistent populations are increasingly important to recovery efforts for the species. A recent study shows that reintroducing frogs from persistent populations into vacant habitat can establish self-sustaining populations with low long-term extinction probability [@knapp2024]. It is critical, therefore, to study the dynamics of these unique populations to gain insights into how they are persisting in the presence of Bd and to assess their value to future conservation efforts.

Among Bd-persistent mountain yellow-legged frog populations, little is known about the ecology and demographics of those in the northern part of their range [@brown2019]. Compared to their relatively well-studied counterparts in the central and southern Sierra Nevada, northern populations are fewer in number, have lower abundances, utilize a more diverse range of habitat, and occupy lower elevations [@interagency2018]. To implement effective conservation measures for mountain yellow-legged frogs in the northern portion of their range, it is vital to better understand the demographics and dynamics of these rare and understudied disease-persistent populations [@brown2020].

In this study, we describe one such population, located in the \_\_\_\_\_\_\_\_\_\_ area of the Tahoe National Forest, California...
